
---
title: "DS 202 – Lab #4 Progress Report"
author: Ash Bhuiyan
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(Lahman)
library(stringr)
```

# Lab report #4 - instructions

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday (after Thanksgiving) to polish things.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 4: Scraping (into) the Hall of Fame

---

### 1. Visualize existing HallOfFame data

```{r visualize-lahman}
hof <- Lahman::HallOfFame
hof %>% 
  ggplot(aes(x = yearID, y = votes/needed*100, group=playerID)) +
  geom_hline(yintercept = 100, colour="grey70") + 
  geom_line() +
  geom_point(aes(colour = inducted), data = hof %>% filter(inducted=="Y")) +
  xlim(2000, 2022) +
  ylab("Percent of votes")
```

### 2. Scrape and clean 2025 Hall of Fame data

```{r scrape-clean-2025}
url <- "https://www.baseball-reference.com/awards/hof_2025.shtml"
raw_tbl <- read_html(url) %>%
  html_element("table") %>%
  html_table(fill = TRUE)

# Promote the first row to column names
names(raw_tbl) <- raw_tbl[1, ]
hof_2025_raw <- raw_tbl[-1, ]

# FIX: Make column names unique
names(hof_2025_raw) <- make.names(names(hof_2025_raw), unique = TRUE)

# Now you can safely filter and mutate
hof_2025_clean <- hof_2025_raw %>%
  filter(!is.na(Name), Name != "", !is.na(Votes), Votes != "") %>%
  distinct(Name, .keep_all = TRUE)

hof_2025 <- hof_2025_clean %>%
  mutate(
    playerID = str_replace_all(Name, " ", ""),
    yearID = 2025L,
    votedBy = "BBWAA",
    ballots = NA_integer_,
    needed = NA_integer_,
    votes = as.integer(Votes),
    # Note: The column for percentage may be called "X.vote" if it was "%vote" originally
    pct_vote = as.numeric(str_replace(X.vote, "%", "")),
    inducted = if_else(pct_vote >= 75, "Y", "N"),
    category = "Player",
    needed_note = NA_character_
  ) %>%
  select(playerID, yearID, votedBy, ballots, needed, votes, pct_vote, inducted, category, needed_note)
updated_hof <- bind_rows(HallOfFame, hof_2025)
```

### 3. Visualize updated Hall of Fame dataset

```{r plot-updated}
updated_hof %>% 
  ggplot(aes(x = yearID, fill = inducted)) +
  geom_bar() +
  xlim(1936, 2025) +
  labs(title = "Hall of Fame Elections, 1936-2025", x = "Year", y = "Number of Players")
```

---

## Ash Bhuiyan's Contribution: Analysis and Summary

### 1. Histogram of 2025 Vote Percentages

```{r hist-2025}
hof_2025 %>%
  ggplot(aes(x = pct_vote)) +
  geom_histogram(binwidth = 5, boundary = 0, closed = "left") +
  scale_x_continuous(breaks = seq(0,100,10)) +
  labs(
    title = "Histogram of 2025 Hall-of-Fame Vote Percentages",
    x = "Vote Percentage",
    y = "Number of Candidates"
  ) +
  geom_vline(xintercept = 75, linetype="dashed", color = "red") +
  annotate("text", x = 80, y = 3, label = "75% inducted", vjust = -0.5)
```

### 2. Summary Statistics for 2025 Ballot

```{r summary-2025}
hof_2025 %>%
  summarize(
    mean_pct   = mean(pct_vote, na.rm = TRUE),
    median_pct = median(pct_vote, na.rm = TRUE),
    sd_pct     = sd(pct_vote, na.rm = TRUE),
    min_pct    = min(pct_vote, na.rm = TRUE),
    max_pct    = max(pct_vote, na.rm = TRUE)
  )
```

| Statistic   | Value   |
|:------------|:-------:|
| **Mean**    | `r round(mean(hof_2025$pct_vote, na.rm=TRUE), 1)`% |
| **Median**  | `r round(median(hof_2025$pct_vote, na.rm=TRUE), 1)`% |
| **Std. Dev**| `r round(sd(hof_2025$pct_vote, na.rm=TRUE), 1)`% |
| **Min**     | `r min(hof_2025$pct_vote, na.rm=TRUE)`% |
| **Max**     | `r max(hof_2025$pct_vote, na.rm=TRUE)`% |

### 3. Reflection

In 2025, the Hall of Fame ballot showed a wide range of support for candidates. The histogram above shows most players received under 50%, with a few standouts above the 75% threshold for induction.

---

*I have completed (analysis/summary/visualization). - Ash Bhuiyan*
